ES Models: Random Forest

### 2025-03-11 docs/old/landscapemetrics_quantile.Rmd
new results

ggbiplot::ggbiplot(wfs_pca_all, groups = wfs_grp$habcat)

@@

ggbiplot::ggbiplot(wfs_pca_all, groups = wfs_grp$habcat, choices = c(2,3))

@@

lsm_pc <- wfs_grp %>% cbind(predict(wfs_pca_all))

>>>>

ggbiplot::ggbiplot(wfs_pca_all, groups = wfs_grp$habcat) +
  egg::theme_article()

@@

ggbiplot::ggbiplot(wfs_pca_all, groups = wfs_grp$habcat, choices = c(2,3)) +
  egg::theme_article()

@@

lsm_pc <- wfs_grp %>% cbind(predict(wfs_pca_all))
saveRDS(lsm_pc, "data/wfs_lsm_pc.rds")

### 2025-03-11 old/landscapemodels_margin.Rmd
new results

>>>>

```{r plot_pol2, include=FALSE}

# Above makes predictions based on population averages for all other variables. Look at predictions with assumed low and high levels of annual agriculture for other variables

# polplot.df_agr5 <- plot_fun(pol.gam1, 
#                        c("aglu_iji", "prn_agr_pland", "aglu_frac_mn", "lu_shdi", "lu_ta"),
#                        wfs_lsm, "ann_agr_pland",
#                        return_data = FALSE,
#                        typical = "mean",
#                        condition = c(ann_agr_pland = 5))

# polplot.df_agr75 <- plot_fun(pol.gam1, 
#                        c("aglu_iji", "prn_agr_pland", "aglu_frac_mn", "lu_shdi", "lu_ta"),
#                        wfs_lsm, "ann_agr_pland",
#                        typical = "mean",
#                        condition = c(ann_agr_pland = 75))

# ggplot(polplot.df[[1]], aes(x, delta_pollinators.margin)) +  
#   geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) + geom_line() +
#   geom_point(data = polplot.df[[2]], 
#              mapping = aes(color = ann_agr_pland), 
#              size = 0.6) + scale_color_continuous(type = "viridis") +
#   facet_wrap(~pred.var, scales = "free_x") +
#   egg::theme_article()

### 2025-03-11 randomforest_kl.Qmd
new results

    wfs <- readRDS("landscape-complexity-main/data/wfs_lsm.rds")

    dat.predictors <-  wfs[, 20:50]

@@

    if(file.exists("ES-RF.RDA")) load("ES-RF.RDA")

@@

        dat <- data.frame(y = wfs$delta_pollinators.margin, dat.predictors)
        pollinators_margin.rf <- ranger(y ~ ., data = dat, importance = "permutation", num.trees = 1000)
        pollinators_margin.p <- data.frame(importance_pvalues(pollinators_margin.rf, method = "altmann", formula = y ~ ., data = dat, num.permutations = 1000))
        rm(dat)

@@

        dat <- data.frame(y = wfs$delta_n_export.margin, dat.predictors)
        n_export_margin.rf <- ranger(y ~ ., data = dat, importance = "permutation", num.trees = 1000)
        n_export_margin.p <- data.frame(importance_pvalues(n_export_margin.rf, method = "altmann", formula = y ~ ., data = dat, num.permutations = 1000))
        rm(dat)

@@

        dat <- data.frame(y = wfs$delta_sed.margin, dat.predictors)
        sed_export_margin.rf <- ranger(y ~ ., data = dat, importance = "permutation", num.trees = 1000)
        sed_export_margin.p <- data.frame(importance_pvalues(sed_export_margin.rf, method = "altmann", formula = y ~ ., data = dat, num.permutations = 1000))
        rm(dat)

@@

        dat <- data.frame(y = wfs$delta_qb.margin, dat.predictors)
        qb_margin.rf <- ranger(y ~ ., data = dat, importance = "permutation", num.trees = 1000)
        qb_margin.p <- data.frame(importance_pvalues(qb_margin.rf, method = "altmann", formula = y ~ ., data = dat, num.permutations = 1000))
        rm(dat)

@@

    save(list = ls(pattern = "^[npqs]"), file = "ES-RF.RDA")

>>>>

    # wfs0 <- readRDS("data/old/wfs_lsm.rds")
      wfs <- readRDS("data/allresults_20250308.rds")

    dat.predictors <-  wfs[, c(81:107,120:123)]

@@

    # if(file.exists("RandomForest/ES-RF.RDA")) load("ES-RF.RDA")

@@

        dat <- data.frame(y = wfs$delta_pollinators_max.margin, dat.predictors)
        pollinators_margin.rf <- ranger(y ~ ., data = dat, importance = "permutation", num.trees = 1000)
        pollinators_margin.p <- data.frame(importance_pvalues(pollinators_margin.rf, method = "altmann", formula = y ~ ., data = dat, num.permutations = 1000))
        rm(dat)

@@

        dat <- data.frame(y = wfs$delta_n_export_max.margin, dat.predictors)
        n_export_margin.rf <- ranger(y ~ ., data = dat, importance = "permutation", num.trees = 1000)
        n_export_margin.p <- data.frame(importance_pvalues(n_export_margin.rf, method = "altmann", formula = y ~ ., data = dat, num.permutations = 1000))
        rm(dat)

@@

        dat <- data.frame(y = wfs$delta_sed_max.margin, dat.predictors)
        sed_export_margin.rf <- ranger(y ~ ., data = dat, importance = "permutation", num.trees = 1000)
        sed_export_margin.p <- data.frame(importance_pvalues(sed_export_margin.rf, method = "altmann", formula = y ~ ., data = dat, num.permutations = 1000))
        rm(dat)

@@

        dat <- data.frame(y = wfs$delta_qb_max.margin, dat.predictors)
        qb_margin.rf <- ranger(y ~ ., data = dat, importance = "permutation", num.trees = 1000)
        qb_margin.p <- data.frame(importance_pvalues(qb_margin.rf, method = "altmann", formula = y ~ ., data = dat, num.permutations = 1000))
        rm(dat)

@@

    save(list = ls(pattern = "^[npqs]"), file = "data/ES-RF.RDA")
